Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2017-11-30 paperIEEE.tex
Shorten RSP section
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@67 db7c848b-7755-4df2-b763-8796797afa52

RSP is an implementation of a reliable multicast protocol over unreliable UDP
transport. RSP behaves similarly to TCP, it provides full reliability and
ordering of the data, and slow receivers will eventually throttle the
sender. This behavior is needed to guarantee delivery of data in all
situations. Pragmatic generic multicast (PGM~\cite{pgm}) provides full ordering,
but slow clients will disconnect from the multicast session instead of
throttling the send rate.

@@

Our RSP implementation uses a separate protocol thread for each RSP group,
which handles all reads and writes on the multicast group. This thread
implements the protocol handling and communicates with the application threads
through thread-safe queues, as shown in \fig{fRSPProto}. These queues contain
datagrams, which are prefixed by a type-specific header. Each connection has a
configurable number of buffers, (1024 by default) of a configurable MTU (1470
bytes default) which are either free for use or in use for reading or writing.

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/rspPackets}
  \caption{\label{fRSPProto}Reliable Stream Protocol}
\end{figure}

A client opens a listening multicast RSP connection. Similar to TCP, this
listener will create one connected RSP connection for each remote member in the
multicast group; that is, on each node $n$ RSP connection instances exist, $1$
for the local listener and $n-1$ for all other nodes. A multicast group
therefore looks like a fully connected set of TCP connections between $n$ nodes
for the user.

Each member of a multicast group has a group-unique 16 bit identifier, which is
randomly selected, announced to the group, and if no collisions are notified by
any other member, confirmed for this connection. This mechanism allows
addressing of connections with a low protocol overhead. The datagrams used in
this exchange contain no payload, and they contain two bytes type (HELLO,
HELLO\_REPLAY, HELLO\_DENY, HELLO\_CONFIRM), two bytes protocol version and two
bytes connection identifier.

The \textsf{\_appBuffers} queue, which is blocking, contains received UDP
datagrams on incoming connections, or empty buffers for writing on outgoing
connections. The \textsf{\_threadBuffers} queue is lock-free and used by the
protocol thread to store empty buffers on incoming connections, and pre-packaged
UDP datagrams on outgoing connections. The application thread pushes buffers
into this queue, consumed by the protocol thread. A datagram is up to 64
kilobytes (UDP limit, configurable) and contains a header (two bytes type, two
bytes size, two bytes source identifier, two bytes sequence number) and its
payload. The sequence number is used for positive (ack) and negative (nack)
acknowledgments and reordering of packets.

The \textsf{\_writeBuffers std::deque} is only used by the protocol
thread. It contains already sent, but not yet acknowledged datagrams. Its
datagrams are resent upon reception of a nack or recycled into
\textsf{\_appBuffers} when acknowledged by all receivers. Similarly, the
\textsf{\_recvBuffers std::deque} contains received out-of-order packets already
received by an incoming connection. Upon each reception it is checked, and
in-order datagrams are placed into \textsf{\_appBuffers} for consumption by the
application thread.

Handling a smooth packet flow is critical for performance. RSP uses active flow
control to advance the byte stream buffered by the implementation. Each incoming
connection actively acknowledges every $n$ (17 by default) packets fully
received. The incoming connections temporarily offset this acknowledgment by
their connection identifier to avoid bursts of acks. Any missed datagram is
actively nack'ed as soon as it is detected. Write connections continuously
retransmit packets from nack datagrams, and advance their
\textsf{\_writeBuffers} upon reception of acks from all readers. Furthermore,
the writer will explicitly request an ack or nack when it runs out of empty
buffers or finishes its write queue. Nack datagrams may contain multiple ranges
of missed datagrams, as UDP implementations often drop multiple packets at
once.

Congestion control is necessary to optimize bandwidth usage. While TCP uses the
well-known AIMD (additive increase, multiplicative decrease) algorithm, we have
chosen a more aggressive congestion control algorithm of additive increase and
decrease. This has proven experimentally to be more optimal: UDP is often
rate-limited by switches; that is, packets are discarded regularly and not
exceptionally. Only slowly backing of the current send rate helps to stay close
to this limit. Furthermore, our RSP traffic is limited to the local subnet,
making cooparation between multiple data stream less of an issue. Send rate
limiting uses a bucket algorithm, where over time the bucket fills with send
credits, from which sends are substracted. If there are no available credits,
the sender sleeps until sufficient credits are available.

>>>>

RSP is an implementation of a reliable multicast protocol over unreliable UDP
transport. RSP behaves similarly to TCP, it provides full reliability and
ordering of the data, and slow receivers will eventually throttle the sender
through a sliding window algorithm. This behavior is needed to guarantee
delivery of data in all situations. Pragmatic generic multicast (PGM~\cite{pgm})
provides full ordering, but slow clients will disconnect from the multicast
session instead of throttling the send rate.

@@

Our RSP implementation uses a separate protocol thread for each RSP group, which
handles all reads and writes on the multicast group. This thread implements the
protocol handling and communicates with the application threads through
thread-safe queues. These queues contain datagrams of the application byte
stream, prefixed by a type-specific header of at most eight bytes. Each
connection has a configurable number of buffers, (1024 by default) of a
configurable MTU (1470 bytes default) which are either free for use or in
transmission.

Handling a smooth packet flow is critical for performance. RSP uses active flow
control to advance the byte stream buffered by the implementation. Each incoming
connection actively acknowledges every $n$ (17 by default) packets fully
received. The incoming connections temporarily offset this acknowledgment by
their connection identifier to avoid bursts of acks. Any missed datagram is
actively nack'ed as soon as it is detected. Write connections continuously
retransmit packets from nack datagrams, and advance their window upon reception
of acks from all readers. Furthermore, the writer will explicitly request an ack
or nack when it runs out of empty buffers or finishes its write queue. Nack
datagrams may contain multiple ranges of missed datagrams, which is motivated by
the observation that UDP implementations often drop multiple packets at once.

Congestion control is necessary to optimize bandwidth usage. While TCP uses the
well-known AIMD (additive increase, multiplicative decrease) algorithm, we have
chosen a more aggressive congestion control algorithm of additive increase and
additive decrease. This has proven experimentally to be more optimal: UDP is
often rate-limited by switches; that is, packets are discarded regularly and not
exceptionally. Only slowly backing of the current send rate helps to stay close
to this limit. Furthermore, our RSP traffic is limited to the local subnet,
making cooparation between multiple data stream less of an issue. Send rate
limiting uses a bucket algorithm, where over time the bucket fills with send
credits, from which sends are substracted. If there are no available credits,
the sender sleeps until sufficient credits are available.
